---
title: "2: Visualization with ggplot"

duedate: "{{< var duedates.lab2 >}}" # change in the main _variables.yml file and it will cascade here and below
toc: true
toc-depth: 1
---

<!-- Note for Ben: Exericse 3 is new here and includes pivot-wider (reshaping) a little census data as I moved that content up from data wrangling. -->

::: {.callout-warning}
## Due Date
This assignment is due on {{< meta duedate >}}.
:::

{{< var blurbs.homeworkdue >}}

 
## A note on exercises
Instead of putting all of your exercises at the end in 1 section, each of the 3 exercises are contained in a callout box like this:

::: {.callout-note}
## Exercise 1 of 3

Do this stuff
:::

So find those in the assignment below. There are a lot of questions clearly labeled "OPTIONAL" as well -- it helps to at least take a look at them, but only those questions in the labeled boxes are required for this lab assignment.

## `tidyverse` and `ggplot2`
We start by assuming that you are familiar (and have installed) `tidyverse` and `ggplot2`. 

## Mappings Link Data to Things You See

```{r, collapse = T, eval = F}
library(tidyverse)
library(ggplot2)
library(gapminder)

gapminder = gapminder::gapminder # dslabs::gapminder is different
head(gapminder)
p <- ggplot(data = gapminder,
            mapping = aes(x = gdpPercap, y = lifeExp))
p + geom_point()
```

Above we've loaded a new dataset and have started to explore a particular relationship. Before putting in this code yourself, try to intuit what *might* be going on in this ggplot code. Any ideas?

Here's a breakdown of everything that happens after the `p<- ggplot()` call:

- `data = gapminder` tells ggplot to use gapminder dataset, so if variable names are mentioned, they should be looked up in gapminder
- `mapping = aes(...)` shows that the mapping is a function call. There is a deeper logic to this that I will disucss below, but it's easiest to simply accept that this is how you write it. Put another way, the `mapping = aes(...)` argument *links variables* to *things you will see* on the plot.
- `aes(x = gdpPercap, y = lifeExp)` maps the GDP data onto `x`, which is a known aesthetic (the x-coordinate) and life expectancy data onto `y`
  - `x` and `y` are  predefined names that are used by `ggplot` and friends

This should all be review for you. Let's do our first exercise:


::: {.callout-note}
## Exercise 1 of 3

Let's use some new data. Assuming you have `ggplot2` loaded using `library(ggplot2)`, you can load up a dataset called `mpg` using `data(mpg)`.

Among the variables in `mpg` are:

- `displ`, a car’s engine size, in litres. Bigger means more powerful.

- `hwy`, a car’s fuel efficiency on the highway, in miles per gallon (mpg). A car with a low fuel efficiency consumes more fuel than a car with a high fuel efficiency when they travel the same distance.

Generate a scatterplot between these two variables. Does it capture the intuitive relationship you expected? What happens if you make a scatterplot of `class` vs `drv`? Why is the plot not useful?
:::



### What do you see?

It turns out there's a reason for doing all of this:

> "The greatest value of a picture is when it forces us to notice what we never expected to see."" — John Tukey

In the plot you made above, one group of points seems to fall outside of the linear trend. These cars have a higher mileage than you might expect. How can you explain these cars?

Let’s hypothesize that the cars are hybrids. One way to test this hypothesis is to look at the class value for each car. The `class` variable of the `mpg` dataset classifies cars into groups such as compact, midsize, and SUV. If the outlying points are hybrids, they should be classified as compact cars or, perhaps, subcompact cars (keep in mind that this data was collected before hybrid trucks and SUVs became popular).

You can add a third variable, like `class`, to a two dimensional scatterplot by mapping it to an aesthetic. An aesthetic is a visual property of the objects in your plot. Aesthetics include things like the size, the shape, or the color of your points. You can display a point (like the one below) in different ways by changing the values of its aesthetic properties. Since we already use the word "**value**" to describe data, let’s use the word "**level**" to describe aesthetic properties. Thus, we are interested in exploring `class` as a level.

You can convey information about your data by mapping the aesthetics in your plot to the variables in your dataset. For example, you can map the colors of your points to the class variable to reveal the class of each car. To map an aesthetic to a variable, associate the name of the aesthetic to the name of the variable inside `aes()`. `ggplot2` will automatically assign a unique level of the aesthetic (here a unique color) to each unique value of the variable, a process known as scaling. `ggplot2` will also add a legend that explains which levels correspond to which values.

::: {.callout-note}
## Exercise 2 of 3

Using your previous scatterplot of `displ` and `hwy`, map the colors of your points to the class variable to reveal the class of each car. What conclusions can we make?

:::


### Back to gapminder::gapminder

Let's explore our previously saved `p` (which used `gapminder::gapminder`) in greater detail. As with Exercise 1, we'll add a *layer*. This says how some data gets turned into concrete visual aspects.

```
p + geom_point()
p + geom_smooth()
```

**Note:** Both of the above geom's use the same mapping, where the x-axis represents `gdpPercap` and the y-axis represents `lifeExp`. You can find this yourself with some ease. But the first one maps the data to individual points, the other one maps it to a smooth line with error ranges.

We get a message that tells us that `geom_smooth()` is using the method = 'gam', so presumably we can use other methods. Let's see if we can figure out which other methods there are.

```
?geom_smooth
p + geom_point() + geom_smooth() + geom_smooth(method = ...) + geom_smooth(method = ...)
p + geom_point() + geom_smooth() + geom_smooth(method = ...) + geom_smooth(method = ..., color = "red")
```

You may start to see why `ggplot2`'s way of breaking up tasks is quite powerful: the geometric objects can all reuse the *same* mapping of data to aesthetics, yet the results are quite different. And if we want later geoms to use different mappings, then we can override them -- but it isn't necessary.

Consider the output we've explored thus far. One potential issue lurking in the data is that most of it is bunched to the left. If we instead used a logarithmic scale, we should be able to spread the data out better.

```
p + geom_point() + geom_smooth(method = "lm") + scale_x_log10()
```
**Try it:** (OPTIONAL) Describe what the `scale_x_log10()` does. Why is it a more evenly distributed cloud of points now? (2-3 sentences.)

Nice. We're starting to get somewhere. But, you might notice that the x-axis now has scientific notation. Let's change that.

```
library(scales)
p + geom_point() +
  geom_smooth(method = "lm") +
  scale_x_log10(labels = scales::dollar)
p + geom_point() +
  geom_smooth(method = "lm") +
  scale_x_log10(labels = scales::...)
```

**Try it:** (OPTIONAL) What does the `dollar()` call do? How can you find other ways of relabeling the scales when using `scale_x_log10()`?

```
?dollar()
```

## The Recipe

1. Tell the `ggplot()` function what our data is.
2. Tell `ggplot()` *what* relationships we want to see. For convenience we will put the results of the first two steps in an object called `p`.
3. Tell `ggplot` *how* we want to see the relationships in our data.
4. Layer on geoms as needed, by adding them on the `p` object one at a time.
5. Use some additional functions to adjust scales, labels, tickmarks, titles.
  - e.g. `scale_`, `labs()`, and `guides()` functions

As you start to run more `R` code, you’re likely to run into problems. Don’t worry — it happens to everyone. I have been writing code in numerous languages for years, and every day I still write code that doesn’t work. Sadly, `R` is particularly persnickity, and its error messages are often opaque.

Start by carefully comparing the code that you’re running to the code in these notes. `R` is extremely picky, and a misplaced character can make all the difference. Make sure that every ( is matched with a ) and every " is paired with another ". Sometimes you’ll run the code and nothing happens. Check the left-hand of your console: if it’s a +, it means that R doesn’t think you’ve typed a complete expression and it’s waiting for you to finish it. In this case, it’s usually easy to start from scratch again by pressing ESCAPE to abort processing the current command.

One common problem when creating ggplot2 graphics is to put the + in the wrong place: it has to come at the end of the line, not the start.



### Mapping Aesthetics vs Setting them

```
p <- ggplot(data = gapminder,
            mapping = aes(x = gdpPercap, y = lifeExp, color = 'yellow'))
p + geom_point() + scale_x_log10()
```

This is interesting (or annoying): the points are not yellow. How can we tell ggplot to draw yellow points?

```
p <- ggplot(data = gapminder,
            mapping = aes(x = gdpPercap, y = lifeExp, ...))
p + geom_point(...) + scale_x_log10()
```

**Try it:** (OPTIONAL) describe in your words what is going on.
One way to avoid such mistakes is to read arguments inside `aes(<property> = <variable>)`as *the property <property> in the graph is determined by the data in <variable>*.

Aesthetics convey information about a variable in the dataset, whereas setting the color of all points to yellow conveys no information about the dataset - it changes the appearance of the plot in a way that is independent of the underlying data.

Remember: `color = 'yellow'` and `aes(color = 'yellow')` are very different, and the second makes usually no sense, as `'yellow'` is treated as *data*.

```
p <- ggplot(data = gapminder,
            mapping = aes(x = gdpPercap, y = lifeExp))
p + geom_point() + geom_smooth(color = "orange", se = FALSE, size = 8, method = "lm") + scale_x_log10()
```

**Try it:** (OPTIONAL)  Write down what all those arguments in `geom_smooth(...)` do.


Coloring by continent:

```
library(scales)
p <- ggplot(data = gapminder,
            mapping = aes(x = gdpPercap, y = lifeExp, color = continent, fill = continent))
p + geom_point()
p + geom_point() + scale_x_log10(labels = dollar)
p + geom_point() + scale_x_log10(labels = dollar) + geom_smooth()
```

**Try it:** (OPTIONAL)  What does `fill = continent` do? What do you think about the match of colors between lines and error bands?



## Tidy up your data, then plot

For Exercise 3, use this [data on MI county-level income, age, population, and yearly commute time](data/milong.csv), which can be loaded using the following line of code:

```{r, eval=F, echo=F, include=F}

mi_county <- get_acs(
  geography = "county",
  state = "MI",
  year = 2023,
  survey = "acs5",
  variables = c(
    population = "B01003_001",
    median_income = "B19013_001",
    median_age = "B01002_001",
    avg_commute_minutes = "B08135_001"
  ),
  output = "wide"
) %>% dplyr::select(GEOID, NAME, ends_with('E'))  %>%
  dplyr::mutate(avg_commute_minutesE = avg_commute_minutesE/populationE) %>%
  pivot_longer(populationE:avg_commute_minutesE, names_to = 'variable', values_to = 'value') %>%
  dplyr::mutate(variable = gsub(x = variable, pattern = 'E$', replacement = '')) %>%
  dplyr::filter(!GEOID %in% c('26083','26095')) # no commute time for these

# write.csv(mi_county, file = file.path('D:/Courses/EC242/data/milong.csv'))

```

```{r, echo=T, eval = F, warning = F, error = F, message = F}
midata = read.csv('https://ec242.netlify.app/data/milong.csv', stringsAsFactors = F)
```



::: {.callout-note}

## Exercise 3 of 3

First, use `pivot_wider` to get the data into tidy form (each observation should be a county). This data is taken from the 2023 US Census American Community Survey, which we'll learn about later on.

Then, categorize each county into "Small" and "Large" based on population. Do the same for age ("Young" and "Old"). A logical cut-point would be the median.

Then, plot the relationship between *two of the variables of your choice* using the geometry of your choice. Use an aesthetic mapping of your choice on the young/old **or** small/large category to illustrate how the relationship may differ across categories. If you're aesthetic mapping young/old to an aesthetic, do not also use it on the X or Y axis (that would be redundant).

Label the axes and the header with clear, easy to understand language. Make sure you label everything, properly capitalize titled, etc. Below your plot, write a few sentences to describe what you've visualized and why.

:::
