---
title: "6: Correlations and Simple Linear Models"
linktitle: "6: Correlations and Simple Linear Models"
duedate: "{{< var duedates.lab6 >}}"
---

::: {.callout-warning}
## Due Date
This assignment is due on **{{< meta duedate >}}**
:::

{{< var blurbs.homeworkdue >}}


## Backstory and Set Up
You have been recently hired to Zillow’s Zestimate product team as a junior analyst. As a part of their regular hazing, they have given you access to a small subset of their historic sales data. Your job is to present some basic predictions for housing values in a small geographic area (Ames, IA) using this historical pricing.

First, let's load the data.

```{r, eval=FALSE}
ameslist  <- read.csv('https://ec242.netlify.app/data/ames.csv',
                      stringsAsFactors = FALSE)
```

Before we proceed, let's note one thing about the (simple) code above. We specify an argument to `read.csv` called `stringsAsFactors`. By default, `read.csv` (the base CSV reading function, which is different from `read_csv`, the `tidyverse` CSV function) turns anything that is a character vector into a factor variable. That's great if you're importing things like state abbreviations. That's not helpful if you're importing character strings that don't have any repitition (e.g. names), or character strings that really should be numeric. We'll handle our character strings our selves, thankyouverymuch.




### Data Exploration and Processing

We are not going to tell you anything about this data. This is intended to replicate a real-world experience that you will all encounter in the (possibly near) future: someone hands you data and you're expected to make sense of it. Fortunately for us, this data is (somewhat) self-contained. We'll first check the variable names to try to divine some information. Recall, we have a handy little function for that:

```{r, eval=FALSE}
names(ameslist)
```

Note that, when doing data exploration, we will sometimes choose to not save our output. This is a judgement call; here we've chosen to merely inspect the variables rather than diving in.

Inspection yields some obvious truths. For example:

| Variable      | Explanation           | Type  |
| ------------- |:-------------:| -----:|
| `ID`      | Unique identifier for each row | `int` |
| `LotArea`    | Size of lot (**units unknown**)      |  `int`|
| `SalePrice` | Sale price of house ($)      |    `int` |

...but we face some not-so-obvious things as well. For example:

| Variable      | Explanation           | Type  |
| ------------- |:-------------:| -----:|
| `LotShape`      | ? Something about the lot| `chr` |
| `MSSubClass`    | ? No clue at all      |  `int`|
| `Condition1` | ? Seems like street info      |    `chr` |

It will be difficult to learn anything about the data that is of type `int` without outside documentation unless it refers to a count of something (e.g. bedrooms). However, we can learn something more about the `chr`-type variables. In order to understand these a little better, we need to review some of the values that each take on. We can use `unique()` to see the unique values it takes. Sometimes, it helps to see how often some value comes up if we're trying to understand a variable's meaning. One handy way of learning this is to use `table()` (which we've seen before) to get a count of the different values a variable can take. **This dataset will have some pernicious `NA`s in it, so when you use `table`, add `useNA = 'always'` as an argument to ensure that we see counts of all values**.

**Try it:** Go through the variables in the dataset and make a note about your interpretation for each. Many will be obvious, but some require additional thought.

Although there are some variables that would be difficult to clean, there are a few that we can address with relative ease. Consider, for instance, the variable `GarageType`. This might not be that important, but, remember, the weather in Ames, IA is pretty crummy---a detached garage might be a dealbreaker for some would-be homebuyers. Let's inspect the values:

```{r, eval=FALSE}
> unique(ameslist$GarageType)
[1] Attchd  Detchd  BuiltIn CarPort <NA> Basment 2Types
```

With this, we could make an informed decision and create a new variable. Let's create `OutdoorGarage` to indicate, say, homes that have any type of garage that requires the homeowner to walk outdoors after parking their car. (For those who aren't familiar with different garage types, a car port is not insulated and is therefore considered outdoors. A detached garage presumably requires that the person walks outside after parking. The three other types are inside the main structure, and `2Types` we can assume includes at least one attached garage of some sort). 

::: {.callout-note}

## EXERCISE 1 of 5

1. Use `case_when` to add a `OutdoorGarage` column to `ameslist` that takes the value of `1` when the house has an outdoor garage, and `0` otherwise. Make sure this is a numeric variable (type `int`). It's up to you to take a stand on what to do with `NA` values. Are those outdoor? Indoors? Do we drop all `NA` values? You often have to decide what to do in cases like this using your knowledge of the context. **Document your reasoning** for how you handle `NA` values.

:::

Generally speaking, this is a persistent issue, and you will spend an extraordinary amount of time dealing with missing data or data that does not encode a variable exactly as you want it. This is expecially true if you deal with real-world data: you will need to learn how to handle `NA`s. There are a number of fixes (as always, Google is your friend) and anything that works is good. But you should spend some time thinking about this and learning at least one approach.

Our goal now is to learn something about correlates between home sale price and the rest of the data. Along the way, you may want to create more variables like `OutdoorGarage` -- for now, make sure those variables are represented as `0` and `1`. We'll work with factor variables more later.




::: {.callout-note}
## EXERCISES 2-5

2. Prune the data to 6-8 of the variables that are `type = int` about which you have some reasonable intuition for what they mean. Choose those that you believe are likely to be correlated with `SalePrice`. This **must** include the variable `SalePrice` and `GrLivArea`. Call this new dataset `Ames`. Produce documentation for this object in the form of a [Markdown table](/resource/markdown/#tables-1) or see [further documentation here](https://www.markdownguide.org/extended-syntax/). This must describe each of the 6-8 preserved variables, the values it can take (e.g., can it be negative?) and your definition of the variable. Counting the variable name, this means your table should have three columns. Markdown tables are entered in the text body, not code chunks, of your .rmd, so your code creating `Ames` will be in a code chunk, and your table will be right after it.


3. Produce a *scatterplot matrix* of the chosen variables[^5]

4. Compute a matrix of correlations between these variables using the function `cor()`. Do the correlations match your prior beliefs? Briefly discuss the correlation between the chosen variables and `SalePrice` and any correlations between these variables.

5. Produce a scatterplot between `SalePrice` and `GrLivArea`. Run a linear model using `lm()` to explore the relationship. Finally, use the `geom_abline()` function to plot the relationship that you've found in the simple linear regression. You'll need to extract the intercept and slope from your `lm` object. See `coef(...)` for information on this.[^6]
    - What is the largest outlier that is above the regression line? Produce the other information about this house.

**(Bonus)** Create a visualization that shows the rise of air conditioning over time in homes in Ames.

:::

[^5]: If you are not familiar with this type of visualization, consult the book (*Introduction to Statistical Learning*), Chapters 2 and 3. 

[^6]: We could also use `geom_smooth(method = 'lm')` to add the regression line, but it's good practice to work with `lm` objects.
