---
title: "Illustrating Classification"
type: docs
weight: 1
editor_options:
  chunk_output_type: console
---

::: {.callout-note}

Today's example will build on material from the previous lecture (earlier this week).

:::


```{r, eval = F, echo=F}
require(readxl)
require(here)
UCI_credit = read_excel(here('data','UCI_credit_file.xls'), skip = 1, col_names=T)
  
UCI_credit = UCI_credit %>%
  dplyr::select(-ID) %>%
  rename(default = `default payment next month`) %>%
  relocate(default) %>%
  dplyr::mutate(SEX = case_when(SEX==1 ~ 'male',
                                SEX==2 ~ 'female'),
                EDUCATION = case_when(EDUCATION==1 ~ 'graduate school',
                                      EDUCATION==2 ~ 'university',
                                      EDUCATION==3 ~ 'high school',
                                      EDUCATION==4 ~ 'others'),
                MARRIAGE = case_when(MARRIAGE==1 ~ 'married',
                                     MARRIAGE==2 ~ 'single',
                                     MARRIAGE==3 ~ 'others') )


write.csv(UCI_credit, here('data','UCI_credit.csv'))

```



## Predicting Defaults

Today, we will work with [UC Irvine's Credit Default Dataset](https://archive.ics.uci.edu/dataset/350/default+of+credit+card+clients). I have made a condensed version of it to work with.

```{r, eval = F, echo = F}
library(ISLR)
library(tibble)
Default = ISLR::Default # NOT USED ANYMORE
```

```{r, echo = F, eval=T}
Default = read_csv(here('data','UCI_credit.csv'))
```

```{r, echo = T, eval = F}
library(tidyverse)
Default = read_csv('https://ec242.netlify.app/data/UCI_credit.csv')
```

### About the data
I've imported the data from the UCI source. Columns `PAY_0` through `PAY_6` show the last 6 months of payment status, where `-2` means "no payment required", `-1` and `0` mean paid on  time, and `1` and above mean 1 month behind, 2 months behind, etc. Bill and Payment amounts as labeled, with 1 being most recent, and 6 being six months ago.




::: {.callout-note}
In our first breakout:

1.Check to make sure the `default` column is a binary indicator for default. Check to make sure there are no surprise `NA` values (there are!). Then, construct any additional variables you might think useful (hint: the `PAY_X` variables don't really have an intuitive numeric interpretation. Should we categorize some of them together?)

2. Explore the data as we did on Tuesday to get an idea of useful predictors.

2. Build a logistic model to predict `default` using any combination of variables and interactions in the data. For now, just use your best judgement for choosing the variables and interactions.

3. Use a Bayes Classifier cutoff of .50 to generate your classifier output.

:::


Let's look at how we did. What variables were most useful in explaining `default`? 


::: {.callout-note}
In our second breakout, we will create a ROC curve manually. To do this

1. Take your model from the first breakout, and using a loop (or `sapply`), step through a large number of possible cutoffs for classification ranging from 0 to 1.

2. For each cutoff, generate a confusion matrix with *accuracy*, *sensitivity* and *specificity*.

3. Combine the cutoff with the *sensitivity* and *specificity* results and make a ROC plot. Use `ggplot` for your plot and map the color aesthetic to the cutoff value.

4. Calculate the AUC (the *area under the curve*). This is a little tricky but can be done with your data.
:::



```{r, eval=F, include=F, echo=F}
Default = na.omit(Default)

mod1 = glm(default ~ . , Default, family = 'binomial')

 pred = ifelse(predict(mod1, type = 'response')>.5, 1,0)
 conf = confusionMatrix(table(predicted = pred, actual = Default$default), positive = '1')
 
  data.frame(Accuracy = conf$overall['Accuracy'],
         Sensitivity = conf$byClass['Sensitivity'],
         Specificity = conf$byClass['Specificity'])


cs = seq(from=.0001, to = .978, length = 1000)

results = lapply(cs, function(cutoff){
  print(cutoff)
  pred = ifelse(predict(mod1, type = 'response')>cutoff, 1, 0)
  conf = confusionMatrix(table(predicted = pred, actual = Default$default), positive = '1')
  
  tibble(Accuracy =    conf$overall['Accuracy'],
         Sensitivity = conf$byClass['Sensitivity'],
         Specificity = conf$byClass['Specificity'])
})

results_compiled = bind_rows(results)

ggplot(results_compiled, aes(x = Specificity, y = Sensitivity)) + geom_line() +
  scale_x_reverse()

sum(results_compiled$Specificity*(1/NROW(results_compiled)))

```
