---
title: "Advanced Visualizations"
draft: false
lastmod: "`r Sys.Date()`"
output:
  blogdown::html_page:
    toc: true
menu:
  example:
    parent: Examples
    weight: 1
type: docs
weight: 1
editor_options:
  chunk_output_type: console
---

```{r, echo=F}
library(tidyverse)
filter <- dplyr::filter
library(emo)

```

<!-- Ben: See answer at very very end and code for solution (not on course materials) -->
<!-- Note: Didn't do the heatmap Try-It as a smaller share of students remembered the SF package. In future, may need to spend more time introducing SF and GIS data? -->

# Data visualization in practice {#gapminder}

In this lecture, we will demonstrate how relatively simple __ggplot2__ code can create insightful and aesthetically pleasing plots. As motivation we will create plots that help us better understand trends in world health and economics. We will implement what we learned in previous sections of the class and learn how to augment the code to perfect the plots. As we go through our case study, we will describe relevant general data visualization principles.

We're going to need the geospatial packages `sf` and the ever-useful `tigris` (for US maps), so install those if you don't already have them.


## Case study: new insights on poverty

Hans Rosling^[https://en.wikipedia.org/wiki/Hans_Rosling] was the co-founder of the Gapminder Foundation^[http://www.gapminder.org/], an organization dedicated to educating the public by using data to dispel common myths about the so-called developing world. The organization uses data to show how actual trends in health and economics contradict the narratives that emanate from sensationalist media coverage of catastrophes, tragedies, and other unfortunate events. As stated in the Gapminder Foundation's website:

> Journalists and lobbyists tell dramatic stories. That’s their job. They tell stories about extraordinary events and unusual people. The piles of dramatic stories pile up in peoples' minds into an over-dramatic worldview and strong negative stress feelings: "The world is getting worse!", "It’s we vs. them!”, “Other people are strange!", "The population just keeps growing!" and "Nobody cares!"

Hans Rosling conveyed actual data-based trends in a dramatic way of his own, using effective data visualization. This section is based on two talks that exemplify this approach to education: [New Insights on Poverty]^[https://www.ted.com/talks/hans_rosling_reveals_new_insights_on_poverty?language=en] and The Best Stats You've Ever Seen^[https://www.ted.com/talks/hans_rosling_shows_the_best_stats_you_ve_ever_seen]. Specifically, in this section, we use data to attempt to answer the following two questions:

1. Is it a fair characterization of today's world to say it is divided into western rich nations and the developing world in Africa, Asia, and Latin America?
2. Has income inequality across countries worsened during the last 40 years?

To answer these questions, we will be using the `gapminder` dataset provided in __dslabs__. This dataset was created using a number of spreadsheets available from the Gapminder Foundation. You can access the table like this:

```{r load libraries, message=FALSE, echo=TRUE, warning = FALSE}
library(tidyverse)
library(dslabs)
library(ggrepel)
library(ggthemes)
gapminder = dslabs::gapminder %>% as_tibble()
```

```{r chunkset, echo=FALSE}
knitr::opts_chunk$set(
  warning = FALSE, error = FALSE, message = FALSE
)
```

### Exploring the Data

Taking an exercise from the _New Insights on Poverty_ video, we start by testing our knowledge regarding differences in child mortality across different countries. For each of the six pairs of countries below, which country do you think had the highest child mortality rates in 2015? Which pairs do you think are most similar?

1. Sri Lanka or Turkey
2. Poland or South Korea
3. Malaysia or Russia
4. Pakistan or Vietnam
5. Thailand or South Africa

When answering these questions without data, the non-European countries are typically picked as having higher child mortality rates: Sri Lanka over Turkey, South Korea over Poland, and Malaysia over Russia. It is also common to assume that countries considered to be part of the developing world: Pakistan, Vietnam, Thailand, and South Africa, have similarly high mortality rates.

To answer these questions __with data__, we can use __tidyverse__ functions. For example, for the first comparison we see that:

```{r, message=FALSE}
dslabs::gapminder %>%
  filter(year == 2015 & country %in% c("Sri Lanka","Turkey")) %>%
  select(country, infant_mortality)
```
Turkey has the higher infant mortality rate.

We can use this code on all comparisons and find the following:

<details class="collapsible-bg">
<summary><strong>Click here to reveal results.</strong></summary>

```{r, echo = FALSE, warning = FALSE}
comp_table <- tibble(comparison = rep(1:5, each = 2),
           country = c("Sri Lanka", "Turkey", "Poland", "South Korea", "Malaysia", "Russia", "Pakistan","Vietnam","Thailand","South Africa"))

tmp <- gapminder %>%
  filter(year == 2015) %>%
  select(country, infant_mortality) %>%
  mutate(country = as.character(country)) ##to match characters to characters

tab <- inner_join(comp_table, tmp, by = "country") %>% select(-comparison)

tmp <- bind_cols(slice(tab,seq(1,9,2)), slice(tab,seq(2,10,2)))
names(tmp) <- c("country", "infant mortality", "country", "infant mortality")
if(knitr::is_html_output()){
  knitr::kable(tmp, "html") %>%
    kableExtra::kable_styling(bootstrap_options = "striped", full_width = FALSE)
} else{
  knitr::kable(tmp, "latex", booktabs = TRUE) %>%
    kableExtra::kable_styling(font_size = 8)
}
```

We see that the European countries on this list have higher child mortality rates: Poland has a higher rate than South Korea, and Russia has a higher rate than Malaysia. We also see that Pakistan has a much higher rate than Vietnam, and South Africa has a much higher rate than Thailand. It turns out that when Hans Rosling gave this quiz to educated groups of people, the average score was less than 2.5 out of 5, worse than what they would have obtained had they guessed randomly. This implies that more than ignorant, we are misinformed. In this chapter we see how data visualization helps inform us by presenting *patterns in the data* that might not be obvious on first glance.

</details>




### Slope charts 
The slopechart is informative when you are comparing variables of the same type, but at different time points and for a relatively small number of comparisons. For example, comparing life expectancy between 2010 and 2015. In this case, we might recommend a _slope chart_.

There is no geometry for slope charts in __ggplot2__, but we can construct one using `geom_line`. We need to do some tinkering to add labels. We'll paste together a character stright with the country name and the starting life expectancy, then do the same with just the later life expectancy for the right side. Below is an example comparing 2010 to 2015 for large western countries: 

```{r slope-plot}
west <- c("Western Europe","Northern Europe","Southern Europe",
          "Northern America","Australia and New Zealand")

dat <- gapminder %>%
  filter(year%in% c(2010, 2015) & region %in% west &
           !is.na(life_expectancy) & population > 10^7) %>%
    mutate(label_first = ifelse(year == 2010, paste0(country, ": ", round(life_expectancy, 1), ' years'), NA),
           label_last = ifelse(year == 2015,  paste0(round(life_expectancy, 1),' years'), NA))

dat %>%
  mutate(location = ifelse(year == 2010, 1, 2),
         location = ifelse(year == 2015 &
                             country %in% c("United Kingdom", "Portugal"),
                           location+0.22, location),
         hjust = ifelse(year == 2010, 1, 0)) %>%
  mutate(year = as.factor(year)) %>%
  ggplot(aes(year, life_expectancy, group = country)) +
  geom_line(aes(color = country), show.legend = FALSE) +
  geom_text_repel(aes(label = label_first, color = country), direction = 'y', nudge_x = -1, seed = 1234, show.legend = FALSE) +
  geom_text_repel(aes(label = label_last, color = country), direction = 'y', nudge_x =  1, seed = 1234, show.legend = FALSE) +
  xlab("") + ylab("Life Expectancy")
```

An advantage of the slope chart is that it permits us to quickly get an idea of changes based on the slope of the lines. Although we are using angle as the visual cue, we also have position to determine the exact values. Comparing the improvements is a bit harder with a scatterplot:


```{r scatter-plot-instead-of-slope, echo=FALSE}
west <- c("Western Europe","Northern Europe","Southern Europe",
          "Northern America","Australia and New Zealand")

dat <- gapminder %>%
  filter(year%in% c(2010, 2015) & region %in% west &
           !is.na(life_expectancy) & population > 10^7)

dat %>%
  mutate(year = paste0("life_expectancy_", year)) %>%
  select(country, year, life_expectancy) %>%
  spread(year, life_expectancy) %>%
  ggplot(aes(x=life_expectancy_2010,y=life_expectancy_2015, label = country)) +
  geom_point() + geom_text_repel() +
  scale_x_continuous(limits=c(78.5, 83)) +
  scale_y_continuous(limits=c(78.5, 83)) +
  geom_abline(lty = 2) +
  xlab("2010") +
  ylab("2015")
```

In the scatterplot, we have followed the principle _use common axes_ since we are comparing these before and after. However, if we have many points, slope charts stop being useful as it becomes hard to see all the lines.

### Bland-Altman plot

Since we are primarily interested in the difference, it makes sense to dedicate one of our axes to it. The Bland-Altman plot, also known as the Tukey mean-difference plot and the MA-plot, shows the difference versus the average:

```{r, bland-altman, echo=T}

dat %>%
  group_by(country) %>%
  filter(year %in% c(2010, 2015)) %>%
  dplyr::summarize(average = mean(life_expectancy),
                   difference = life_expectancy[year==2015]-life_expectancy[year==2010]) %>%
  ggplot(aes(average, difference, label = country)) +
  geom_point() +
  geom_text_repel() +
  geom_abline(lty = 2) +
  xlab("Average of 2010 and 2015") +
  ylab("Difference between 2015 and 2010")
```

Here, by simply looking at the y-axis, we quickly see which countries have shown the most improvement. We also get an idea of the overall value from the x-axis. You already made a similar Altman plot in an earlier problem set, so we'll move on.






### Bump charts

Finally, we can make a bump chart that shows changes in rankings over time. We'll look at fertility in South Asia. First we need to calculate a new variable that shows the rank of each country within each year. We can do this if we group by year and then use the `rank()` function to rank countries by the `fertility` column.

```{r filter-bump-data}
sa_fe <- gapminder %>%
  filter(region == "Southern Asia") %>%
  filter(year >= 2004, year < 2015) %>%
  group_by(year) %>%
  mutate(rank = rank(fertility))
```

We then plot this with points and lines, reversing the y-axis so 1 is at the top:

```{r make-bump-plot}
ggplot(sa_fe, aes(x = year, y = rank, color = country)) +
  geom_line() +
  geom_point() +
  scale_y_reverse(breaks = 1:8) +
  labs(x = "Year", y = "Rank", color = "Country")
```

Iran holds the number 1 spot, while Sri Lanka dropped from 2 to 6, and Bangladesh increased from 4 to 2.

As with the slopegraph, there are 8 different colors in the legend and it's hard to line them all up with the different lines, so we can plot the text directly instead. We'll use `geom_text()` again. We don't need to repel anything, since the text should fit in each row just fine. We need to change the `data` argument in `geom_text()` though and filter the data to only include one year, otherwise we'll get labels on every point, which is excessive. We can also adjust the theme and colors to make it cleaner.

```{r bump-plot-fancier, echo = TRUE, fig.width=7, fig.height=4.2, warnings = F}
bumpplot <- ggplot(sa_fe, aes(x = year, y = rank, color = country)) +
  geom_line(size = 2) +
  geom_point(size = 4) +
  geom_text(data = sa_fe %>% dplyr::filter(year==2004) %>% arrange(rank),
            aes(label = country, x = 2003), fontface = "bold", angle = 45) +
 geom_text(data = sa_fe %>% dplyr::filter(year==2014) %>% arrange(rank),
            aes(label = country, x = 2015), fontface = "bold", angle = 45) + 
  guides(color = 'none') + # another way of turning off legend
  scale_y_reverse(breaks = 1:8) +
  scale_x_continuous(breaks = 2004:2014) +
  scale_color_viridis_d(option = "C", begin = 0.2, end = 0.9) +
  labs(x = NULL, y = "Rank") +
  theme(plot.margin = unit(c(0.5, 0.5, 0.5, 0.5),  "inches")) 

bumpplot
```




### Themes
We can go a little further towards a clean, easy-to-read data visualization by using __themes__ in our plots. Themes allow us to set a particular range of plot settings in one command, and let us further tweak things like fonts, background colors, and much more. We've used them in passing a few times without highlighting them, but we'll discuss them here.

A pre-constructed set of instructions for making a visual theme can be had by using a theme's `ggplot` function. Let's look at two of my favorites.

`theme_bw()` uses the black-and-white theme, which is helpful in making a nice, clean plot:

```{r, echo = T,  fig.width=7, fig.height=5.2}
bumpplot + theme_bw()
```

The background shading is gone, which gives the plot a nice, crisp feel. It adds the black outline around the plot, but doesn't mess with the colors in the plot.

Here's `theme_minimal()`

```{r, echo = T,  fig.width=7, fig.height=5.2}
bumpplot + theme_minimal()
```

Themes can alter things in the plot as well. If we really want to strip it down and remove the Y-axis (which is rarely a good idea, but in a bump chart, it makes sense):

```{r, echo = T,  fig.width=7, fig.height=5.2}
bumpplot + theme_void()
```

Now **that's** clean!

In our opening unit, we had a plot that was styled after the plots in the magazine, *The Economist*. That's a theme (in the `ggthemes` package that we loaded at the top)!

```{r, echo = T,  fig.width=7, fig.height=5.2}
bumpplot + theme_economist()
```

Themes affect some of the plot elements that we haven't gotten much into (like length of axis ticks and the color of the panel grid behind the plot). We can use a theme, then make further changes to the theme. We won't go into a lot of detail, but here's an example. Use the `?theme` to learn more about what you can change. Half the challenge is finding the right term for the thing you want to tweak! Theme changes occur in code order, so you can update a pre-set theme with your own details:

```{r bump-plot-fancierest, echo = TRUE, fig.width=7, fig.height=5.2}
bumpplot +   theme_bw() + theme(strip.text = element_text(face = "bold"),
                   plot.title = element_text(face = "bold"),
                   axis.text.x = element_text(angle = 45, hjust = 1),
                   panel.grid.major.y = element_blank(), # turn off all of the Y grid
                   panel.grid.minor.y = element_blank(),
                   panel.grid.minor.x = element_blank()) # turn off small x grid
```

### Theme elements
The above code looks a little odd in that asking it to leave out the minor and major Y grid, and the major X grid, required `element_blank()`, a function! Since a "blank" part (or a solid color part) might entail a lot of particular things, the ggplot function is necessary here to take care of all the particular details. 

Relatedly, if we wanted to change the background color of some panel (the plotted area), then `element_rect()` would be used because the panel is a rectangle. The theme argument would be `panel.background =element_rect(fill="red")` if we wanted to make it red (and hideous)

```{r bump-plot-fancierestest, echo = TRUE, fig.width=7, fig.height=5.2}
bumpplot +   theme_bw() + theme(strip.text = element_text(face = "bold"),
                                panel.background  = element_rect(fill='red'),
                   plot.title = element_text(face = "bold"),
                   axis.text.x = element_text(angle = 45, hjust = 1),
                   panel.grid.major.y = element_blank(), # turn off all of the Y grid
                   panel.grid.minor.y = element_blank(),
                   panel.grid.minor.x = element_blank()) # turn off small x grid
```

We can also set the `strip.background` for the "strip" rectangles that label our `facet_wrap` and `facet_grid`sections,and `plot.background` for the area behind the plot panel.

In the above example, we also set the `element_text` function for the axis text and specified a rotation (see the years are at 45 degrees) and a horizontal-adjustment so that they center over the line. `element_text` is a theme element that controls how the text looks (including font, face, color, etc). This is particularly useful when you have appropriate axis labels that might be long, like the `scales::label_comma` [from earlier this week](/content/Week_04/04a.html/#labelerFunctions). We can avoid scrunching our labels by using `element_text` to set a new angle (45 or 90).




## Case study: vaccines and infectious diseases {#vaccines}

Vaccines have helped save millions of lives. In the 19th century, before herd immunization was achieved through vaccination programs, deaths from infectious diseases, such as smallpox and polio, were common. However, today vaccination programs have become somewhat controversial despite all the scientific evidence for their importance.

The controversy started with a paper^[http://www.thelancet.com/journals/lancet/article/PIIS0140-6736(97)11096-0/abstract] published in 1988 and led by Andrew Wakefield claiming
there was a link between the administration of the measles, mumps, and rubella (MMR) vaccine and the appearance of autism and bowel disease.
Despite much scientific evidence contradicting this finding, sensationalist media reports and fear-mongering from conspiracy theorists led parts of the public into believing that vaccines were harmful. As a result, many parents ceased to vaccinate their children. This dangerous practice can be potentially disastrous given that the Centers for Disease Control (CDC) estimates that vaccinations will prevent more than 21 million hospitalizations and 732,000 deaths among children born in the last 20 years (see Benefits from Immunization during the Vaccines for Children Program Era — United States, 1994-2013, MMWR^[https://www.cdc.gov/mmwr/preview/mmwrhtml/mm6316a4.htm]).
The 1988 paper has since been retracted and Andrew Wakefield was eventually "struck off the UK medical register, with a statement identifying deliberate falsification in the research published in The Lancet, and was thereby barred from practicing medicine in the UK." (source: Wikipedia^[https://en.wikipedia.org/wiki/Andrew_Wakefield]). Yet misconceptions persist, in part due to self-proclaimed activists who continue to disseminate misinformation about vaccines.

Effective communication of data is a strong antidote to misinformation and fear-mongering. Earlier we used an example provided by a Wall Street Journal article^[http://graphics.wsj.com/infectious-diseases-and-vaccines/] showing data related to the impact of vaccines on battling infectious diseases.  Here we reconstruct that example.

The data used for these plots were collected, organized, and distributed by the Tycho Project^[http://www.tycho.pitt.edu/]. They include weekly reported counts for seven diseases from 1928 to 2011, from all fifty states. The yearly totals are helpfully included in the __dslabs__ package:

```{r message=FALSE, warning=FALSE, echo = TRUE}
library(RColorBrewer)
data(us_contagious_diseases)
names(us_contagious_diseases)
```

We create a temporary object `dat` that stores only the measles data, includes a per 100,000 rate, orders states by average value of disease and removes Alaska and Hawaii since they only became states in the late 1950s. Note that there is a `weeks_reporting` column that tells us for how many weeks of the year data was reported. We have to adjust for that value when computing the rate.

```{r}
the_disease <- "Measles"
dat <- us_contagious_diseases %>%
  filter(!state%in%c("Hawaii","Alaska") & disease == the_disease) %>%
  mutate(rate = count / population * 10000 * 52 / weeks_reporting) %>%
  mutate(state = reorder(state, rate))
```

We can now easily plot disease rates per year. Here are the measles data from California:

```{r california-measles-time-series}
dat %>% filter(state == "California" & !is.na(rate)) %>%
  ggplot(aes(year, rate)) +
  geom_line() +
  ylab("Cases per 10,000")  +
  geom_vline(xintercept=1963, col = "blue")
```

We add a vertical line at 1963 since this is when the vaccine was introduced [Control, Centers for Disease; Prevention (2014). CDC health information for international travel 2014 (the yellow book). p. 250. ISBN 9780199948505].

Now can we show data for all states in one plot? We have three variables to show: year, state, and rate. In the WSJ figure, they use the x-axis for year, the y-axis for state, and color hue to represent rates. However, the color scale they use, which goes from yellow to blue to green to orange to red, can be improved.

In our example, we want to use a sequential palette since there is no meaningful center, just low and high rates.

We use the geometry `geom_tile` to tile the region with colors representing disease rates. We use a square root transformation to avoid having the really high counts dominate the plot. Notice that missing values are shown in grey. Note that once a disease was pretty much eradicated, some states stopped reporting cases all together. This is why we see so much grey after 1980.

```{r vaccines-plot, out.width="100%", fig.height=5}
dat %>% 
  mutate(state = factor(state, levels = rev(levels(state)[order(levels(state))]))) %>%
  ggplot(aes(year, state, fill = rate)) +
  geom_tile(color = "grey50") +
  scale_x_continuous(expand=c(0,0)) +
  scale_fill_gradientn(colors = brewer.pal(9, "Reds"), trans = "sqrt") +
  geom_vline(xintercept=1963, col = "blue") +
  theme_minimal() +
  theme(panel.grid = element_blank(),
        legend.position="bottom",
        text = element_text(size = 8)) +
  ggtitle(the_disease) +
  ylab("") + xlab("")
```

This plot makes a very striking argument for the contribution of vaccines. However, one limitation of this plot is that it uses color to represent quantity, which we earlier explained makes it harder to know exactly how high values are going. Position and lengths are better cues. If we are willing to lose state information, we can make a version of the plot that shows the values with position. We can also show the average for the US, which we compute like this:

```{r}
avg <- us_contagious_diseases %>%
  filter(disease==the_disease) %>% group_by(year) %>%
  summarize(us_rate = sum(count, na.rm = TRUE) /
              sum(population, na.rm = TRUE) * 10000)
```

Now to make the plot we simply use the `geom_line` geometry:
```{r time-series-vaccines-plot}
dat %>%
  filter(!is.na(rate)) %>%
    ggplot() +
  geom_line(aes(year, rate, group = state),  color = "grey50",
            show.legend = FALSE, alpha = 0.2, size = 1) +
  geom_line(mapping = aes(year, us_rate),  data = avg, size = 1) +
  scale_y_continuous(trans = "sqrt", breaks = c(5, 25, 125, 300)) +
  ggtitle("Cases per 10,000 by state") +
  xlab("") + ylab("") +
  geom_text(data = data.frame(x = 1955, y = 50),
            mapping = aes(x, y, label="US average"),
            color="black") +
  geom_vline(xintercept=1963, col = "blue")
```

In theory, we could use color to represent the categorical value state, but it is hard to pick 50 distinct colors.


## Saving your plot

In RMarkdown, we display the plot as part of the chunk output. For your group project (or for future use) you'll likely need to save your plot as well. The function `ggsave` is a good way of outputting your plot. 

```
ggsave(myPlot, file = 'myPlot.png', width = 7, height = 5, dpi=300)
```
File type will be discerned from the suffix on the file, and the height and width are in inches. The `dpi` argument sets the pixes-per-inch. Generally, screens display at around 150 (though this is rapidly increasing, it used to be 72 dpi), and print is usually 300 or better. A width of 5 inches at 300 dpi will render an image 1,500 pixels across.

The dpi and file size may change the *relative* font size, so you might have to experiment with different sizes and dpi values for exporting.


::: {.callout-note}

## TRY IT

1. Reproduce the heatmap plot we previously made but for smallpox. For this plot, do not include years in which cases were not reported in 10 or more weeks.

2. Now reproduce the time series plot we previously made, but this time following the instructions of the previous question for smallpox.

3. For the state of California, make a time series plot showing rates for all diseases. Include only years with 10 or more weeks reporting. Use a different color for each disease.

4. Now do the same for the rates for the US. Hint: compute the US rate by using summarize: the total divided by total population.

:::




# Geospatial visualization

**Note: I assume you have met the pre-requisites for this course, which includes PLS202. PLS202 covers the basics of geospatial in R using the `sf` package. If you are unfamiliar or need a review, I have a short lesson (that used to be part of our course material before PLS202 started including it) located [in a resource here](/resource/geospatialbasics.qmd)**

In the measles case, it might be nice to examine measles rates on a map. Maybe there's something regional about vaccine rollout? Or maybe we think there might be spillovers between states (measles is incredibly highly contagious).

We'll need `tigris` with accesses US census TIGER/line files which define nearly all federally-recognized boundaries, from states to counties and territories, to US census tracts and blocks. Today we'll work only with US states, since that's the *level of granularity* at which our measles data is reported (that is, we don't see e.g. county-leve measles cases).

```{r, echo=T, warning = F, message = F, error = F}
library(sf)
library(tigris)
```

## `geom_sf()`

Lucky for us, a `sf` object is all we need to make a nice visual map. The `sf` object is just a tidyverse `tbl` that has a special column, the `geometry` column, that gives the geospatial location and shape (polygon or point or line) that corresponds to that data.

Let's map the `rate` we used in the heatmap above. Since we have many years of observations, let's use 1960 (before the vaccine) and 1965 (2 years after the vaccine).

```{r, echo=T, collapse = T, message = F, error = F, warning = F}

states = tigris::states(progress_bar = F)
# progress_bar = F just keeps the function from putting out a progress bar, which takes up output space
# probably use it on your projects, but not necessary for interactive.

head(states)
ggplot(states) + geom_sf()
```

We get a lot of information besides the `geometry` column, including the census `REGION` and `DIVISION`, plus the `STATEFP`, which is the "state FIPS code" for each state. These are unique numbers for each state that are used in federal data. We'll use them more in the future.

We also get all the territories, which extends the map window way out the South Pacific. Let's use some tricks to make this better. First off, US territories have FIPS codes of 60 and above, so we can filter to get the 50 US States + DC. Then, we can shift the geometry to include AK and HI (not to scale).

```{r, echo = T, message = F, error = F, warning = F, collapse = T}
states = tigris::states(cb = T, progress_bar = F) %>%
  dplyr::filter(STATEFP<60) %>%
  tigris::shift_geometry(position = 'below') %>%
  dplyr::select(STATEFP, STATEFP, STUSPS, NAME)

head(states)

ggplot(states) + geom_sf() + theme_minimal()
```

Just a note: we selected only 4 columns, but `R` (rather, the `sf` package) kept the `geometry` column automatically! That's a feature of `sf` -- it won't let you drop that important column unless you specifically tell it to.

Now we need to `join` our measles data to our map. We're covering `join` in the next lecture, so we'll keep it simple here. Note that we're joining `dat` to the `states` object using a key of `NAME`.

```{r, echo = T, message = F, error = F, warning = F}
measles.to.merge.1960 = dat %>% dplyr::filter(year==1960) %>% dplyr::select(rate, NAME = state)
# Note: I re-named state-->NAME so it matches our states column name
# 
measles.to.merge.1965 = dat %>% dplyr::filter(year==1965) %>% dplyr::select(rate, NAME = state)


states.measles.1960 = left_join(x = states,
                                y = measles.to.merge.1960,
                                by = 'NAME')

states.measles.1965 = left_join(x = states,
                                y = measles.to.merge.1965,
                                by = 'NAME')

head(states.measles.1960)

```

Alright, we now have two geospatial objects, one for 1960, one for 1965. Let's make the plots and take a peek at 1960:

```{r, echo = T, message = F, error = F, warning = F}

p.1960 = ggplot(states.measles.1960, aes(fill = rate)) +
  geom_sf() +
  scale_fill_continuous(type = 'viridis') +
  labs(fill = 'Measles Rate', title = 'Measles - 1960') +
  theme_void() +# really get rid of the extra stuff
  theme(legend.position = "bottom") 

p.1960
```

A couple things to note in this code:

- We use the `fill` aesthetic mapping, so the state borders still have the default for `color`
- Kansas, Alaska, and Hawaii are `NA` for 1960
- We set the theme to `theme_void()` and *then* use another `theme()` function to move the legend. Think of it as overwriting the previous theme.



Finally, let's put the years side-by-side:

```{r, echo = T, message = F, error = F, warning = F}
library(gridExtra)
library(patchwork)

p.1965 = ggplot(states.measles.1965, aes(fill = rate)) +
  geom_sf() +
  scale_fill_continuous(type = 'viridis') +
  labs(fill = 'Measles Rate', title = 'Measels - 1965') +
  theme_void() +# really get rid of the extra stuff
  theme(legend.position = "bottom")

p.1960 + p.1965 # this is patchwork

```


:::{.callout-note}
## TRY IT

There's something wrong (or, at the very least, unhelpful) about the finished measles plot above. Can anyone guess what it is?

And can anyone fix it?
:::

```{r, echo = F, message = F, error = F, warning = F, eval = F}
## The problem is the guide scale doesn't line up.
## 
p.1960b = ggplot(states.measles.1960, aes(fill = rate)) +
  geom_sf() +
  scale_fill_continuous(type = 'viridis', limits = c(min(c(states.measles.1960$rate, states.measles.1965$rate), na.rm=T),
                                                     max(c(states.measles.1960$rate, states.measles.1965$rate), na.rm=T))) +
  labs(fill = 'Measles Rate', title = 'Measles - 1960') +
  theme_void() +# really get rid of the extra stuff
  theme(legend.position = "bottom") 

p.1965b = ggplot(states.measles.1965, aes(fill = rate)) +
  geom_sf() +
  scale_fill_continuous(type = 'viridis', limits = c(min(c(states.measles.1960$rate, states.measles.1965$rate), na.rm=T),
                                                     max(c(states.measles.1960$rate, states.measles.1965$rate), na.rm=T))) +
  labs(fill = 'Measles Rate', title = 'Measels - 1965') +
  theme_void() +# really get rid of the extra stuff
  theme(legend.position = "bottom")

p.1960b + p.1965b + plot_layout(guides = "collect") &
  theme(legend.position='bottom')# this is patchwork
```


