---
title: "Lab No. X"
author: "Your Name Here"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
 pdf_document:
     pandoc_args: ["--extract-media", "."]
---


```{r setup, include=TRUE, echo = FALSE}
# Leave this chunk as it is. It sets defaults for subsequent chunks.
# The "echo" option tells Rmarkdown whether or not it should show the code in the chunk.
# Setting echo = TRUE outputs both the code and the results
knitr::opts_chunk$set(echo = TRUE,
                     collapse = TRUE,
                     warning = FALSE, message = FALSE, error = FALSE,
                     fig.align = 'center')
```





## 1. Text of the question copied here (replace this with the question text)

### a. If there is a subquestion, use this with three ###'s at the beginning of the line, and a space afterwards


```{r}
# code goes here, replacing this placeholder code
a = 'hello world'
print(a)
```

Your answer to the first question. Replace this text with your own. Use markdown language for formatting (**bold** and *italics*) (or [see here for more](https://docs.github.com/en/get-started/writing-on-github/getting-started-with-writing-and-formatting-on-github/basic-writing-and-formatting-syntax#styling-text)). 


## 2. Text of next question copied here. etc.

```{r}
# Your code for 1B here.
```

Your answer to the second question/subquestion goes here. If it's only asking you to code and there is no blue "Question", then you can just say "no question" here.



You're on your own for the structure here on out. When you make a new section with the two or three hashtags (pound signs), they have to be at the beginning of the line - no spaces before them. Check your work. Delete this sentence before rendering your final PDF.
