---
title: "Illustrating Classification"
type: docs
weight: 1
editor_options:
  chunk_output_type: console
---

::: {.callout-note}

Today's example will build on material from the previous lecture (earlier this week).

:::


```{r, eval = F, echo=F}
require(readxl)
require(here)
UCI_credit = read_excel(here('data','UCI_credit_file.xls'), skip = 1, col_names=T)
  
UCI_credit = UCI_credit %>%
  dplyr::select(-ID) %>%
  rename(default = `default payment next month`) %>%
  relocate(default) %>%
  dplyr::mutate(SEX = case_when(SEX==1 ~ 'male',
                                SEX==2 ~ 'female'),
                EDUCATION = case_when(EDUCATION==1 ~ 'graduate school',
                                      EDUCATION==2 ~ 'university',
                                      EDUCATION==3 ~ 'high school',
                                      EDUCATION==4 ~ 'others'),
                MARRIAGE = case_when(MARRIAGE==1 ~ 'married',
                                     MARRIAGE==2 ~ 'single',
                                     MARRIAGE==3 ~ 'others') )


write.csv(UCI_credit, here('data','UCI_credit.csv'))

```



## Predicting Defaults

Today, we will work with [UC Irvine's Credit Default Dataset](https://archive.ics.uci.edu/dataset/350/default+of+credit+card+clients). I have made a condensed version of it to work with.

```{r, eval = F, echo = F}
library(ISLR)
library(tibble)
Default = ISLR::Default # NOT USED ANYMORE
```

```{r, echo = F, eval=T}
Default = read_csv(here('data','UCI_credit.csv'))
```

```{r, echo = T, eval = F}
library(tidyverse)
library(caret)
Default = read_csv('https://ec242.netlify.app/data/UCI_credit.csv')
```

### About the data
I've imported the data from the UCI source. Note that the first column is just a leftover row number from the export -- `read_csv` will call it `...1` -- so drop it before you fit anything with `default ~ .`. Columns `PAY_0`, `PAY_2`, `PAY_3`, `PAY_4`, `PAY_5`, and `PAY_6` show the last 6 months of payment status (there is no `PAY_1` -- that's how UCI shipped it; `PAY_0` is the most recent month), where `-2` means "no payment required", `-1` and `0` mean paid on time, and `1` and above mean 1 month behind, 2 months behind, etc. Bill and Payment amounts as labeled, with 1 being most recent, and 6 being six months ago. All the dollar figures, `LIMIT_BAL` included, are New Taiwan dollars -- so a `LIMIT_BAL` of 20000 is not a \$20,000 credit line.




::: {.callout-note}
In our first breakout:

1. Check to make sure the `default` column is a binary indicator for default. Check to make sure there are no surprise `NA` values (there are!). Then, construct any additional variables you might think useful (hint: the `PAY_X` variables don't really have an intuitive numeric interpretation. Should we categorize some of them together?)

2. Explore the data as we did on Tuesday to get an idea of useful predictors.

3. Build a logistic model to predict `default` using any combination of variables and interactions in the data. For now, just use your best judgement for choosing the variables and interactions.

4. Use a Bayes Classifier cutoff of .50 to generate your classifier output.

:::


Let's look at how we did. What variables were most useful in explaining `default`? 


::: {.callout-note}
In our second breakout, we will create a ROC curve manually. To do this

1. Take your model from the first breakout, and using a loop (or `sapply`), step through a large number of possible cutoffs for classification ranging from 0 to 1. Careful at the ends: a cutoff below every fitted probability (or above every one) predicts a single class, so `table()` gives you a one-row table and `confusionMatrix()` won't accept it -- the same problem we ran into on Tuesday with the all-"No" classifier. Keep your cutoffs inside the range of your predicted probabilities.

2. For each cutoff, generate a confusion matrix with *accuracy*, *sensitivity* and *specificity*.

3. Combine the cutoff with the *sensitivity* and *specificity* results and make a ROC plot. Use `ggplot` for your plot and map the color aesthetic to the cutoff value.

4. Calculate the AUC (the *area under the curve*). This is a little tricky but can be done with your data.
:::



```{r, eval=F, include=F, echo=F}
Default = na.omit(Default) %>% dplyr::select(-1)

mod1 = glm(default ~ . , Default, family = 'binomial')

 pred = ifelse(predict(mod1, type = 'response')>.5, 1,0)
 conf = confusionMatrix(table(predicted = pred, actual = Default$default), positive = '1')
 
  data.frame(Accuracy = conf$overall['Accuracy'],
         Sensitivity = conf$byClass['Sensitivity'],
         Specificity = conf$byClass['Specificity'])


cs = seq(from=.0001, to = .978, length = 1000)

results = lapply(cs, function(cutoff){
  print(cutoff)
  pred = ifelse(predict(mod1, type = 'response')>cutoff, 1, 0)
  conf = confusionMatrix(table(predicted = pred, actual = Default$default), positive = '1')
  
  tibble(cutoff = cutoff,
         Accuracy =    conf$overall['Accuracy'],
         Sensitivity = conf$byClass['Sensitivity'],
         Specificity = conf$byClass['Specificity'])
})

results_compiled = bind_rows(results)

ggplot(results_compiled, aes(x = Specificity, y = Sensitivity, color = cutoff)) + geom_line() +
  scale_x_reverse()

# AUC = area under Sensitivity plotted against FPR = 1 - Specificity
fpr = 1 - results_compiled$Specificity
tpr = results_compiled$Sensitivity
o = order(fpr)
fpr = fpr[o]; tpr = tpr[o]
sum(diff(fpr) * (head(tpr, -1) + tail(tpr, -1)) / 2)

```
