---
title: "Visualizations in Practice"
lastmod: "`r Sys.Date()`"
output:
  blogdown::html_page:
    toc: true
menu:
  example:
    parent: Examples
    weight: 1
type: docs
weight: 1
editor_options:
  chunk_output_type: console
execute:
  freeze: auto  # re-render only when source changes
---

<!-- PostFall2025 note: Week 6a ALSO talks about distributions and viz, but we didn't get to it in F25. Didn't seem to matter. Make sure the material in F25-06a on distributions is in this week so 6a doesn't need it in the future (can use the time to have them code their own simulation probably? Prisoners Loop?) -->

<!-- Note to Ben (S26 onward): I've moved this entire unit on histograms, densities, the normal distribution, and basic geometries (barplot, etc.) to a resource and replaced it with 4a (condensing down). -->

<!-- after-F25: the try-its at the end didn't really work. Need to make it clear that the regions can be used to collapse the data further. Probably need a hint to start by calculating the polling"error". DONE -->



## Readings
 
- This page.  


### Guiding Questions

- Today we're mostly learning some technical aspects of `ggplot`. 
- Why are we covering the material this way? (Good question. There's an answer!)



### Prep
Load up our `murders` data

```{r, warning = FALSE, error = FALSE, message = FALSE}
library(dslabs)
library(ggplot2)
library(dplyr)
data(murders)

p <- ggplot(data = murders, aes(x = population, y = total, label = abb))
```


# Scales and transformations

### Log transformations
In our first visualization lecture, we re-scaled our population by 10^6 (millions), but still had a lot of variation because some states are tiny and some are huge. Sometimes, we want to have one (or both) of our axes scaled non-linearly. For instance, if we wanted to have our x-axis be in log base 10, then each major tick would represent a factor of 10 over the last. This is not the default, so this change needs to be added through a _scales_ layer. A quick look at the cheat sheet reveals the `scale_x_continuous` function lets us control the behavior of scales. We use them like this:


```{r ggplot-example-9}
p + geom_point(size = 3) +
  geom_text(nudge_x = 0.05) +
  scale_x_continuous(trans = "log10") +
  scale_y_continuous(trans = "log10")
```

A couple of things here: adding things like `scale_x_continuous(...)` operates on the whole plot. In some cases, order matters, but it doesn't here, so we can throw `scale_x_continuous` anywhere. Because we have altered the whole plot's scale to be in the log-scale now, the _nudge_ must be made smaller. It is in log-base-10 units. Using `?scale_x_continuous` brings us to the help for both `scale_x_continuous` and `scale_y_continuous`, which shows us the options for transformations `transform = ...` (you'll see `trans = ...` in older code, including the chunk above -- it still works, but it's deprecated)

This particular transformation is so common that __ggplot2__ provides the specialized functions `scale_x_log10` and `scale_y_log10` which "inherit" (take the place of) the `scale_x_continuous` functions but have log base 10 as default. We can use these to rewrite the code like this:

```{r, echo=TRUE}
p + geom_point(size = 3) +
  geom_text(nudge_x = 0.05) +
  scale_x_log10() +
  scale_y_log10()
```

This can make a plot much easier to read, though one has to be sure to pay attention to the values on the axes. Plotting anything with very large outliers will almost always be better if done in log-scale. Adding the scale layer is an easy way to fix this.

We can also use one of many built-in transformations. Of note: **reverse** just inverts the scale, which can be helpful, **log** uses the natural log, **sqrt** takes the square root (dropping anything with a negative value), **reciprocal** takes 1/x. If your x-axis is in a date format, you can also scale to **hms** (hour-minute-second) or **date**.


### Transforming data vs. transforming using `scale_...`
We could simply take the log of population and log of total in the call and we'd get something very similar. Note that we had to override the aesthetic mapping set in `p` in each of the geometries:

```{r, eval=TRUE}
p + geom_point(aes(x = log(population, base=10), y = log(total, base=10)), size = 3) +
  geom_text(aes(x = log(population, base=10), y = log(total, base=10)), nudge_x = 0.05) 
```

This avoids using `scale_x_continuous` or its child function `scale_x_log10`. One advantage to using `scale_x...` is that the axes are correctly labeled. When we transform the data directly, the axis labels only show the transformed values, so 10,000,000 becomes 7.0. This could be confusing! We could update the axis labels to say "total murders (log base 10)" and "total population (log base 10)", but that's cumbersome. Using `scale_x...` is a lot more refined and easy.

# Axis labels, legends, and titles
But let's say we did want to re-name our x-axis label. Or maybe we don't like that the variable column name is lower-case "p". 

As with many things in `ggplot`, there are many ways to get the same result. We'll go over one way of changing titles and labels, but know that there are many more.

## Changing axis titles
We'll use the `labs(...)` annotation layer to do this, which is pretty straightforward. `?labs` shows us what we can change, and while it looks pretty basic, the real meat is in the `...` argument, which the help says is "A list of new name-value pairs". This means we can re-define the label on anything that is an aesthetic mapping. X and Y are aesthetic mappings, so...

```{r, echo=TRUE}
p + geom_point(size = 3) +
  geom_text(nudge_x = 0.05) +
  scale_x_log10() +
  scale_y_log10() + 
  labs(x = 'Population', y = 'Total murders')
```

Now, let's use an aesthetic mapping that generates a legend, like `color`, and see what `labs` renames:


```{r, echo=TRUE}
p + geom_point(aes(color = region), size = 3) +
  geom_text(nudge_x = 0.05) +
  scale_x_log10() +
  scale_y_log10() + 
  labs(x = 'Population', y = 'Total murders', color = 'US Region')
```
We can rename the aesthetic mapping-relevant label using `labs`. Even if there are multiple mapped aesthetics:

```{r, echo=TRUE}
p + geom_point(aes(color = region, size = total/population)) +
  geom_text(nudge_x = 0.05) +
  scale_x_log10() +
  scale_y_log10() + 
  labs(x = 'Population', y = 'Total murders', color = 'US Region', size = 'Murder rate')
```

## Plot Titles
In `?labs`, we also see some things that look like titles and captions. We can include those:

```{r, echo=TRUE}
p + geom_point(aes(color = region, size = total/population)) +
  geom_text(nudge_x = 0.05) +
  scale_x_log10() +
  scale_y_log10() + 
  labs(x = 'Population', y = 'Total murders', color = 'US Region', size = 'Murder rate',
       title = 'This is a title', subtitle = 'This is a subtitle', caption = 'This is a caption', tag = 'This is a tag')
```

Now that you know how, **always label your plots with at least a title and have meaningful axis and legend labels**.

## Axis ticks
In addition to the axis labels, we may want to format or change the axis tick labels (like "1e+06" above) or even where the tick marks and lines are drawn. If we don't specify anything, the axis labels and tick marks are drawn as best as `ggplot` can do, but we can change this. This might be especially useful if our data has some meaningful cutoffs that aren't found by the default, or we just don't like where the marks fall or  how they are labeled. This is easy to fix with `ggplot`.

To change the tick mark labels, we have to set the tick mark locations. Then we can set a label for each tick mark. Let's go back to our `murders` data and take control of the Y-axis tick marks. We'll use `scale_y_continuous` to tell R *where* to put the breaks (`breaks = `) and *what to label* the breaks. We have to give it one label for every break. Let's say we just want a line at the 500's and let's say we want to (absurdly) use written numerics for each of the Y-axis lines. Since `scale_y_log10` **inherits from** `scale_y_continuous`, we can just use that and add the breaks and labels:

```{r, echo=TRUE}
p + geom_point(aes(color = region), size = 3) +
  geom_text(nudge_x = .05) +
  scale_x_log10() +
  scale_y_log10(breaks = c(0,50, 100, 500,1000,1500), 
                     labels = c('Zero','Fifty','One hundred','Five hundred','One thousand','Fifteen hundred')) +
  labs(x = 'Population', y = 'Total murders', color = 'US Region')
```
We have manually set both the location and the label for the y-axis. Note that R filled in the in-between "minor" tick lines, but we can take those out. Since we are setting the location of the lines, we can do anything we want:

```{r, echo=TRUE}
p + geom_point(aes(color = region), size = 3) +
  geom_text(nudge_x = .05) +
  scale_x_log10() +
  scale_y_log10(breaks = c(0,50, 100, 721, 1000,1500), 
                     labels = c('Zero','Fifty','One hundred','Seven hundred twenty one','One thousand','Fifteen hundred'),
                     minor_breaks = NULL) +
  labs(x = 'Population', y = 'Total murders', color = 'US Region')
```

So we can now define where axis tick lines should lie and how they should be labeled.

### Flexible axis tick labels
Anyone else annoyed by the scientific `1e+06`-style x-axis labels? If you're not familiar, `1e+06` means `1 * 10^6` or `1,000,000`. Of course, we could manually make the labels (and the breaks) like we did above, but that might be tedious, and what if we have different ranges of data? We'd end up forcing tick marks in areas that are far out of the support of the data (off the plot).

We can also use a `labeler` function. The most common ones are in the `scales` package, so you'll want to install that and load it up using `library(scales)`. Each labeler is a function that takes as its first argument a vector of numbers (or some expected format, like dates) and returns a formatted character vector of labels.

So the following will happen:

- `scales::label_comma`
```{r}
numbers.vector = c(20000, 50000, 100000, 500000, 1000000) # 20k to 1M

scales::label_comma(numbers.vector, big.mark = ',', decimal.mark = '.')

```

Wha? That looks like the output...is a function?

**Yes. Yes it is.** The function returns....a function. And that function takes the number and returns the labels:

```{r}

scalefun = scales::label_comma(big.mark = ',', decimal.mark = '.')

scalefun(numbers.vector)

```

Now we have the appropriate format! Here are some others:

```{r}
scalefun.percent = scales::label_percent(accuracy = .001, scale=100)

scalefun.percent(murders$total[1:5]/murders$population[1:5])

scalefun.dollar = scales::label_dollar(accuracy = 1, scale=1)

scalefun.dollar(numbers.vector)

```

Note that we could set the `accuracy` and `scale` arguments in `scales::label_percent` and `scales::label_dollar`. The `accuracy` argument determines rounding. We don't need cents when we have dollar values up to 1M, but we do need fractions of a percent when we use the murder rate (not multiplied by 100,000).

## Using labeler functions {#labelerFunctions}
The above was just an illustration of how the labelers work. To use it in practice, we need only give the function name to ggplot:

```{r, echo=TRUE}
p + geom_point(aes(color = region), size = 3) +
  geom_text(nudge_x = .05) +
  scale_x_log10(labels= scales::label_comma()) +
  scale_y_log10(breaks = c(0,50, 100, 721, 1000,1500), 
                     labels = scales::label_comma(accuracy=.01),
                     minor_breaks = NULL) +
  labs(x = 'Population', y = 'Total murders', color = 'US Region')
```

We don't need the decimals on the y-axis tick labels, it's just there for illustration.

Sometimes, our axis tick labels are still too long. When numbers get long, we usually have standard ways of representing them (like k, M, G, and T for kilo-, mega-, giga- and terabytes). Labeler functions can make this easy for us.

We can use the `labels = label_comma(scale_cut = cut_short_scale())` argument to shorten the display and use the common suffix.

```{r, echo=T}
p + geom_point(aes(color = region), size = 3) +
  geom_text(nudge_x = .05) +
  scale_x_log10(labels= scales::label_comma(scale_cut = scales::cut_short_scale())) +
  scale_y_log10(labels= scales::label_comma(scale_cut = scales::cut_short_scale())) +
  labs(x = 'Population', y = 'Total murders', color = 'US Region')
```

Of course, we could also transform the `Population` and `total` variables, dividing by 1,000,000 or 1,000, then add "Thousands" or "Millions" to the label. But this way is cleaner. Also, notice on the Y-axis that the suffix "K" is only used when necessary -- we don't have ".01K" on there.

There is also a `cut_si()` function that takes as an input a scientific unit (like "g" for grams). It'll use the full scientific notation for larger quantities, so `cut_si("g")` would have "g", then "kg", then "Mg" where needed on the axis.

There are lots of other neat tricks for axis labels on the tidyverse [scales vignette](https://tidyverse.org/blog/2022/04/scales-1-2-0/)

## Axis tick label orientation

While the `label_comma` kindly leaves room for us on the x-axis breaks (by eliminating some of the tick labels), we can also alter the orientation of the labels. This occurs in the `theme()` call since it adjusts the layout or styling of the plot. Here, I've put in a bunch of breaks at 1 million to 40 million, and then asked `ggplot` to rotate and adjust the x-axis tick labels. `hjust` and `vjust` move the horizontal and vertical justification of the label -- where the text sits relative to its tick mark. `angle` changes the orientation. Careful, though: justification follows the rotated text, so at `angle = 90` it's `hjust` that slides the label up and down and `vjust` that centers it on its tick.

```{r, echo=TRUE}
p + geom_point(aes(color = region), size = 3) +
  geom_text(nudge_x = .05) +
  scale_x_log10(breaks = c(1,2,3,4,5,10,15,20,25,30, 40)*1000000, labels= scales::label_comma()) +
  scale_y_log10(breaks = c(0,50, 100, 721, 1000,1500), 
                     labels = scales::label_comma(accuracy=1),
                     minor_breaks = NULL) +
  labs(x = 'Population', y = 'Total murders', color = 'US Region') + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5))
```

Note that rotating the axis text unfortunately doesn't allow the labeler to add more breaks by default (leaving `breaks = ...` empty wouldn't add more breaks to the x axis). Also, `?element_text` will let you see the other things you can change (color, font family)



## Additional geometries
Let's say we are happy with our axis tick locations, but we want to add a single additional line. Maybe we want to divide at 1,000,000 population (a vertical line at 1,000,000) because we think those over 1,000,000 are somehow different, and we want to call attention to the data around that point. As a more general example, if we were to plot, say, car accidents by age, we would maybe want to label age 21, when people can legally purchase alcohol (and subsequently cause car accidents).

`geom_vline` lets us add a single vertical line (without aesthetic mappings). If we look at `?geom_vline` we see that it requires one aesthetic: `xintercept`. It also takes aesthetics like color and size, and introduces the `linetype` aesthetic:

```{r, echo=TRUE}
p + geom_point(aes(color = region), size = 3) +
  geom_text(nudge_x = .05) +
  geom_vline(aes(xintercept = 1000000), col = 'red', linewidth = 2, linetype = 2) +
  scale_x_log10() +
  scale_y_log10(breaks = c(0,50, 100, 721, 1000,1500), 
                     labels = c('Zero','Fifty','One hundred','Seven hundred twenty one','One thousand','Fifteen hundred'),
                     minor_breaks = NULL) +
  labs(x = 'Population', y = 'Total murders', color = 'US Region')
```

Combining geometries is as easy as adding the layers with `+`.

### `geom_line`
For a good old line plot, we use the line geometry at `geom_line`. The help for `?geom_line` tells us that we need an x and a y aesthetic (much like `geom_point`). Since our `murders` data isn't really suited to a line graph, we'll use a daily stock price. We'll get this using `tidyquant`, which pulls stock prices from Yahoo Finance and maintains the "tidy" format. You'll need to `install.packages('tidyquant')` before you run this the first time.

```{r getStock, echo=T, warning = FALSE, error = FALSE, message = FALSE}
library(tidyquant)
AAPL = tq_get("AAPL", from = '2009-01-01', to = '2021-08-01', get = 'stock.prices')
head(AAPL)
```

Now, we can plot a line graph of the Apple closing stock price over the requested date range. We want this to be a time series, so the x-axis will be the date and the y-axis will be the closing price.

```{r geomPath, echo=T}
ggplot(AAPL, aes(x = date, y = close)) +
  geom_line() +
  labs(x = 'Date', y = 'Closing price', title = 'Apple stock price')
```

In `geom_line`, R will automatically sort on the x-variable. If you don't want this, then `geom_path` will use whatever order the data is in. Either way, if you have multiple observations for the same value on the x-axis, then you'll get something pretty messy because R will try to connect, in some order, all the points. Let's see an example with two stocks:

```{r, echo=T}
AAPLNFLX = tq_get(c("AAPL","NFLX"), from = '2021-01-01', to = '2021-08-01', get = 'stock.prices')
ggplot(AAPLNFLX, aes(x = date, y = close)) +
  geom_line() +
  labs(x = 'Date', y = 'Closing price', title = 'Apple and Netflix stock price')
```

That looks kinda strange. That's because, for every date, we have two values - the NFLX and the AAPL value, so each day has a vertical line drawn between the two prices. This is nonsense, especially since what we want to see is the history of NFLX and AAPL over time.

Aesthetics to the rescue! Remember, when we use an aesthetic mapping, we are able to separate out data by things like color or linetype. Let's use color as the aesthetic here, and map it to the stock ticker:

```{r, echo=T}
AAPLNFLX = tq_get(c("AAPL","NFLX"), from = '2021-01-01', to = '2021-08-01', get = 'stock.prices')
ggplot(AAPLNFLX, aes(x = date, y = close, color = symbol)) +
  geom_line() +
  labs(x = 'Date', y = 'Closing price', title = 'Apple and Netflix stock price')
```
Well there we go! We can now see each stock price over time, with a convenient legend. Later on, we'll learn how to change the color palette. If we don't necessarily want a different color but we do want to separate the lines, we can use the `group` aesthetic.

```{r, echo=T}
AAPLNFLX = tq_get(c("AAPL","NFLX"), from = '2021-01-01', to = '2021-08-01', get = 'stock.prices')
ggplot(AAPLNFLX, aes(x = date, y = close, group = symbol)) +
  geom_line() + 
  labs(x = 'Date', y = 'Closing price', title = 'Apple and Netflix stock price')
```

Similar result as mapping `color`, but without the color difference (which makes it rather hard to tell what you're looking at). But if we add labels using `geom_label`, we'll get one label for every point, which will be overwhelming. The solution? Use some filtered data so that there is only one point for each label. But that means replacing the `data` in `ggplot`. Here's how.



# Using different data with different geometries

Just as we can use different aesthetic mappings on each geometry, we can use different *data* entirely. This is useful when we want one geometry to have one set of data (like the stock prices above), but another geometry to only have a subset of the data. Why would we want that? Well, we'd like to label just *one* part of each of the lines in our plot, right? That means we want to label a *subset* of the stock data.

To replace data in a geometry, we just need to specify the `data = ` argument separately:

```{r, echo=T}
ggplot(AAPLNFLX, aes(x = date, y = close, group = symbol)) +
  geom_line() +
  geom_label(data = AAPLNFLX %>% group_by(symbol) %>% slice(100),
             aes(label = symbol),
             nudge_y = 20) + 
  labs(x = 'Date', y = 'Closing price', title = 'Apple and Netflix stock price')
```

In `geom_label`, we specified we wanted the 100th observation from each symbol to be the label location. Then, we nudged it up along y by 20 so that it's clear of the line.

R also has a very useful `ggrepel` package that gives us `geom_label_repel` which takes care of the nudging for us, even in complicated situations (lots of points, lines, etc.). It does a decent job here of moving the label to a point where it doesn't cover a lot of data.

```{r, echo=T, message = F, warning = F, error = F}
library(ggrepel)
ggplot(AAPLNFLX, aes(x = date, y = close, group = symbol)) +
  geom_line() +
  geom_label_repel(data = AAPLNFLX %>% group_by(symbol) %>% slice(100),
             aes(label = symbol)) + 
  labs(x = 'Date', y = 'Closing price', title = 'Apple and Netflix stock price')
```

Now, we don't lose a lot of space to a legend, and we haven't had to use color to separate the stock symbols.


## Summarizing data and including in ggplot
Often, we'll want to have lines or highlights in places that are derived from the data. In the above example from `murders`, we added a line at an arbitrary x-value (population) of 1,000,000. Often, though, **we'll want to show some summary of the data** on the plot. That also requires using a second dataset (derived from the first).

Let's say we want to show the average state population for each region. We'll need to use two different data objects, one for the original, one for the summary. Create the summary first:

```{r, echo=T}
murders.summary = murders %>%
  group_by(region) %>%
  dplyr::summarize(meanpop = mean(population),
                   meantot = mean(total),
                   meanrate = 100000*sum(total)/sum(population))

print(murders.summary)
```

Then, use both `murders` and `murders.summary` in the plot:

```{r, echo=T}
p + geom_point(aes(color = region), size = 3) +
  geom_text(nudge_x = .05) +
  geom_vline(data = murders.summary, 
             aes(xintercept = meanpop, color=region), linewidth = 2, linetype = 1, alpha=.5) +
  scale_x_log10() +
  scale_y_log10() +
  labs(x = 'Population', y = 'Total murders', color = 'US Region')
```

Note that we get the same color mapping -- this isn't necessarily automatic. Rather, it's because `region` is a factor with *the exact same levels* in both `murders` and `murders.summary`, and `ggplot` assigns colors in level order -- not in the order the rows happen to show up. If the levels didn't match, we'd have to set them using `factor` and overwriting our old `region` column:

```{r, echo=T}
murders.summary = murders %>%
  group_by(region) %>%
  dplyr::summarize(meanpop = mean(population),
                   meantot = mean(total),
                   meanrate = 100000*sum(total)/sum(population)) %>%
  dplyr::mutate(region = factor(region, levels = levels(murders$region)))

print(murders.summary)
```

Perhaps we would like to summarize the regional average Population and Murders by adding representative points. Since we're adding new data, we have to remove the `label=abb` part of the `aes` that inherits to each geom, which we do with `label=NULL` in the new point geometry:

```{r}
p + geom_point(aes(color = region), size = 3) +
  geom_text(nudge_x = .05) +
  geom_point(data = murders.summary, aes(x = meanpop, y = meantot, color = region, label = NULL), size =12, shape=13) +
  scale_x_log10() +
  scale_y_log10() +
  labs(x = 'Population', y = 'Total murders', color = 'US Region')

```
<!-- ### Multiple geometries -->
<!-- Since this section is about adding geometries, we *can* combine points and lines. Since lines connect points, it will look like a giant connect-the-dots.  -->

<!-- ```{r, echo=T} -->
<!-- library(ggrepel) -->
<!-- ggplot(AAPLNFLX, aes(x = date, y = close, group = symbol)) + -->
<!--   geom_line() + -->
<!--   geom_point() +  -->
<!--   geom_label_repel(data = AAPLNFLX %>% group_by(symbol) %>% slice(100), -->
<!--              aes(label = symbol)) +  -->
<!--   labs(x = 'Date', y = 'Closing price', title = 'Apple and Netflix stock price') -->
<!-- ``` -->


::: {.callout-note}
## Try it!


#### In-Class Activity: Comparing 2008 Electoral Votes and Polling Data. 

In this activity, we will use `R` to visualize and compare the actual electoral votes from the 2008 presidential election with the polling data prior to the election. Specifically, we will explore how accurately the polls predicted the election results. Yes, I know this is an older election... that's kind of the point. This isn't a politics class. 

**Setup:**

You'll need to download two CSV files and put them somewhere on your computer (or upload it to RStudio.cloud if you've gone that direction)---preferably in a folder named `data` in your project folder. You can download the data from the link below:

- [<i class="fas fa-file-csv"></i> `pres08.csv`](/data/pres08.csv)
- [<i class="fas fa-file-csv"></i> `polls08.csv`](/data/polls08.csv)

**Steps Prior to Analysis:** 

1. Load the Data

```{r, echo=TRUE, message=FALSE, eval=FALSE}
# Load the necessary libraries
library(tidyverse)
electoral_votes <- read_csv("path_to/pres08.csv")
polling_data <- read_csv("path_to/polls08.csv") %>%
  dplyr::mutate(middate = as.Date(middate)) # we'll tackle dates later
```

Note that you'll have to replace `path_to` with the actual path to the files on your computer.

2. Data Preparation

We need to aggregate the polling data to get an average poll result for each state. We will then merge this with the electoral votes data.

```{r, echo=TRUE, message=FALSE, eval=FALSE}
# Aggregating polling data
avg_polls <- polling_data %>%
  group_by(state) %>%
  summarise(Obama_avg = mean(Obama), McCain_avg = mean(McCain))

# Merging with electoral votes data
combined_data <- merge(electoral_votes, avg_polls, by = "state")
```

Then, we need to generate a variable that represents the discrepancy between results (`Obama` and `McCain`) and the polling average (`Obama_avg`). Use `mutate()` to do so.

Finally, we want something to cut the data -- below, we have some definitions for region. Use those (perhaps with `case_when()`?) to load in the state-to-region mapping.

```{r, eval=FALSE}
# Define the regions
Northeast = c('CT', 'ME', 'MA', 'NH', 'RI', 'VT', 'NJ', 'NY', 'PA')
Midwest =  c('IL', 'IN', 'MI', 'OH', 'WI', 'IA', 'KS', 'MN', 'MO', 'NE', 'ND', 'SD')
South = c('DE', 'FL', 'GA', 'MD', 'NC', 'SC', 'VA', 'DC', 'WV', 'AL', 'KY', 'MS', 'TN', 'AR', 'LA', 'OK', 'TX')
West = c('AZ', 'CO', 'ID', 'MT', 'NV', 'NM', 'UT', 'WY', 'AK', 'CA', 'HI', 'OR', 'WA')
```



3. Complete Your Analysis 

You are tasked with two different things:

A: Explore and visualize the discrepancies between polls and actual results using `combined_data`. How do you want to go about representing the data? How many observations do you have per state? What about per region?

You might find the region definitions above to be useful for this (hint: use them with `case_when`).


```{r, eval=F, echo=F}
ggplot(combined_data, aes(x = region, y = Obama_avg - Obama )) + geom_boxplot()
```

B: Visualize the relationship between the actual results and the poll results *over time* for the following states: `c('MI','WI','IA','FL','NC','GA')`. *Does this relationship change over time?* Or, put another way, *were early polls or later polls more accurate*? To do this, you'll need to merge the final `Obama` and `McCain` results from `electoral_votes` into the original `polling_data`. Both contain `state` columns!

**Make sure you label your axes, legends, and other elements correctly.** You may work in groups of up to three.

```{r, eval=F, echo=F}
pd = polling_data %>%
  dplyr::filter(state %in% c('MI','WI','IA','FL','NC','GA')) %>%
  left_join(electoral_votes %>% dplyr::select(state, ObamaActual = Obama, McCainActual = McCain), by = 'state') %>%
  dplyr::mutate(Obama_diff = Obama - ObamaActual)

ggplot(pd, aes(x = middate, y = Obama_diff, col = state)) + geom_line() + facet_wrap(~ state)

```


Use {{< var url.zoom >}} to share. Happy visualizing!

:::




